---
title: "RandomRandevouz"
author: "Alex Alvarez"
date: "2024-04-08"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Lab: Random Rendezvous**

Suppose two people (let’s call them Kareem and Kathleen) agree to meet for lunch at a certain restaurant. But these are both important people who can’t be sure when they’ll be able to get away, so both arrival times are random variables, independent of each other. 

If each person agrees to wait exactly 15 minutes for the other before giving up and leaving, what is the probability that the two of them actually meet?

## **Part 1: Uniform**

First suppose that the arrival time for each person, in minutes after noon, is uniformly distributed between 0 and 60. Also assume that the two arrival times are independent of each other. We will first use a simulation to investigate what happens when the random process of their lunch meeting is repeated for a large number of days.


The R code at the end of this assignment simulates $N$ repetitions of this random process and approximates the probability that the two people successfully meet.


a) Run the R code for 10,000 repetitions, with a wait time of 15 minutes. Produce and submit a scatterplot of the pairs of arrival times. Comment on what the scatterplot reveals, and report the approximate probability that Kareem and Kathleen meet.


```{r}
# User must first enter values for number of repetitions (N) AND how long each person agrees to wait
N = 10000
wait = 15

# Generate random arrival times
time1 = runif(N, 0, 60)
time2 = runif(N, 0, 60)

# Calculate difference AND absolute difference in arrival times
diff = time1 - time2
absdiff = abs(diff)

# Produce graphs of these two variables
hist(diff)
hist(absdiff)

# Determine whether or not meeting occurs
meet = (absdiff < wait)



# Produce scatter plot of arrival times with different symbols for meet/don't meet

plot(time1, time2, type='n')
points(time1[meet=='TRUE'], time2[meet=='TRUE'], pch = 20, col="green")
points(time1[meet=='FALSE'], time2[meet=='FALSE'], pch = 20, col="red")


# Calculate approx probability
probmeet = sum(meet)/N
cat("Probability of meeting: ", probmeet)
```
### Comment: 
x axis is the time katheen gets there, y axis is the time kareem gets there.


b) Increase the number of repetitions to 1,000,000 days. Report the approximate probability that Kareem and Kathleen meet. 
- Because both distributions are uniform and independent of each other, exact probabilities can be calculated by determining the area of the region of interest as a fraction of the total area of the 60×60 square.

```{r}
# User must first enter values for number of repititions (N) AND how long each person agrees to wait
N = 1000000
wait = 15

# Generate random arrival times
time1 = runif(N, 0, 60)
time2 = runif(N, 0, 60)

# Calculate difference AND absolute difference in arrival times
diff = time1 - time2
absdiff = abs(diff)

# Produce graphs of these two variables
hist(diff)
hist(absdiff)

# Determine whether or not meeting occurs
meet = (absdiff < wait)



# Produce scatter plot of arrival times with different symbols for meet/don't meet

plot(time1, time2, type='n')
points(time1[meet=='TRUE'], time2[meet=='TRUE'], pch = 20, col="green")
points(time1[meet=='FALSE'], time2[meet=='FALSE'], pch = 20, col="red")


# Calculate approx probability
probmeet = sum(meet)/N
cat("Probability of meeting: ", probmeet)
```



c) Draw (by hand is fine) a 60×60 square to represent the sample space of all possible pairs of arrival times. Label one of the axes for Kareem’s arrival time and the other for Kathleen’s arrival time. Then shade in the region of the square corresponding to the event that they arrive within 15 minutes of each other and therefore successfully meet.



d) Use geometry to determine the area of the region in which Kareem and Kathleen successfully meet. Then divide by the total area of the square to determine the probability that they meet. Is the actual probability within the error bound of the approximate probability from your simulation analysis?

$$
\text{Total area: 60x60= 3600}\\
\text{Total area of red: 45x45 = 2025}\\
\text{Area of green: 3600 - 2025 = 1575}\\
\text{1575/3600= 0.4375}

$$
# It is in the error bound

e) Now let `m` represent the number of minutes that both people agree to wait, where `m` can be any real number between 0 and 60. Use geometry to express the probability that they successfully meet as a function of m. Also produce (and submit) a graph of this probability as a function of m, and comment on its behavior.
$$
\text{Probability of meeting}\\
\begin{aligned}
f(m) &= 1 - \frac{(60-m)^2}{60^2} \\
&= 1 - \frac{60^2-120m+m^2}{60^2} \\
\end{aligned}

$$

```{r}
probmeet <- function(m){
  return (1 - ((60-m)**2/60**2))
}

fun(15)
m = seq(0,60,1)
p = probmeet(m)
plot(m, p, type='l')
```


```{r}
# User must first enter values for number of repititions (N) AND how long each person agrees to wait
N = 10000
wait = 50

# Generate random arrival times
time1 = runif(N, 0, 60)
time2 = runif(N, 0, 60)

# Calculate difference AND absolute difference in arrival times
diff = time1 - time2
absdiff = abs(diff)

# Produce graphs of these two variables
hist(diff)
hist(absdiff)

# Determine whether or not meeting occurs
meet = (absdiff < wait)



# Produce scatter plot of arrival times with different symbols for meet/don't meet

plot(time1, time2, type='n')
points(time1[meet=='TRUE'], time2[meet=='TRUE'], pch = 20, col="green")
points(time1[meet=='FALSE'], time2[meet=='FALSE'], pch = 20, col="red")


# Calculate approx probability
probmeet = sum(meet)/N
cat("Probability of meeting: ", probmeet)
```

f) Determine how long each person would have to agree to wait in order for this probability (of successfully meeting the other person) to be at least .5. Then determine how long each person would have to agree to wait in order for this probability to be at least .9.
```{r}
getTime <- function(p){
  return (60*(1-sqrt(1-p)))
}

getTime(0.9)
```

Through algebraic manipulation, we will see that in order to get 17.57359313 min for a 50% CHANGE

for a 90% change, it will be 41.0 minutes




---
---



## **Part 2: Normal**
Now suppose that each person’s arrival time, in minutes after noon, follows a normal distribution with mean 30 and standard deviation 10. Continue to assume that they arrive independently of each other and that they agree to wait for 15 minutes.


g) Modify the R code to simulate this random process. [Hint: The command to simulate random data from a normal distribution is: `rnorm(number_of_repetitions, mean, sd)`.] 
- Run the R code for 10,000 repetitions, with a wait time of 15 minutes. 

```{r}
# User must first enter values for number of repititions (N) AND how long each person agrees to wait
N = 1000000
mu = 30
sd = 10
wait = 15

# Generate random arrival times
time1 = rnorm(N, mu, sd)
time2 = rnorm(N, mu, sd)

# Calculate difference AND absolute difference in arrival times
diff = time1 - time2
diff = time1 - time2
cat("Probability of meeting: ", probmeet)
absdiff = abs(diff)

# Produce graphs of these two variables
hist(diff)
hist(absdiff)

# Determine whether or not meeting occurs
meet = (absdiff < wait)



# Produce scatter plot of arrival times with different symbols for meet/don't meet

plot(time1, time2, type='n')
points(time1[meet=='TRUE'], time2[meet=='TRUE'], pch = 20, col="green")
points(time1[meet=='FALSE'], time2[meet=='FALSE'], pch = 20, col="red")


# Calculate approx probability
probmeet = sum(meet)/N
cat("Probability of meeting: ", probmeet)
```


- Produce and submit a scatterplot of the pairs of arrival times. 
- Comment on how this scatterplot differs from the previous one. Also report the approximate probability that Kareem and Kathleen meet. The exact probability that they meet can be determined from a normal distribution, using the result about linear combinations of normal distributions.

### **Comment**
This distribution is different since it no longer takes up the entirety of the square (non uniform) 


h) Report the probability distribution of the difference (not absolute difference) in the arrival times of Kareem and Kathleen. 
[Hint: You might let $T_j$ represent Kareem’s arrival time and $T_k$ represent Kathleen’s arrival time, both in minutes after noon. 
Use what you know about normal distributions to specify the probability distribution of the difference D = Tj – Tk.]
$$

\begin{aligned}
T_j \sim N(30, 10^2) \\
T_k \sim N(30, 10^2) \\
D = T_j - T_k \\
\end{aligned}
\\
$$
we know he linear combination of Normal random variables is a normal random variable
$$
\begin{aligned}
E(T_j - T_k) &= E(T_j) - E(T_k) \\
 &= 30 - 30 \\
 &= 0 \\
\end{aligned}
$$

The Joint variance is:
$$
\begin{aligned}
Var(D) &= Var(T_j) + Var(T_k) \\
 &= 10^2 + 10^2 \\
 &= 200 \\
\end{aligned}
$$
Therefore,
$$
D = T_j - T_k \sim N(0,200)
$$

i) Use appropriate normal probability calculations to determine the probability that the two people successfully meet. Also report the values of the appropriate z-scores. [Hint: First express the probability that they successfully meet in terms of the random variable D.]
Our variance is 200, the standard deviation is then $\sqrt{200}$

$$
\text{Probability the two people meet:}\\
P(-15 < D < 15)
$$
Z Score for $D = -15$
$$
\begin{aligned}
Z &= \frac{-15 - 0}{\sqrt{200}}\\
&= -1.06\\
\end{aligned}
$$
```{r}
x <- -15
μ <- 0  # mean
σ <- sqrt(200)  # standard deviation

# Calculate z-score
z <- (x - μ) / σ

# CDF CALCULATIONS
Φ <- pnorm(z)  # also, percentile of 'x'

# PDF calculation
ϕ <- dnorm(z)

cat("z-score of", x, ": ", z, "\n")
cat("pdf : (f(x) = y val of bell curve at x) = ", ϕ, "\n")
cat("cdf : Pr( X <",x, ") =", Φ, "\n")     # probability below x
cat("1 - cdf: Pr(",x, " < X ) =", (1-Φ), "\n\n")   # probability above X

# Calculate percent of data above a value 'x'
```
Z Score for $D = 15$
$$
\begin{aligned}
Z &= \frac{15 - 0}{\sqrt{200}}\\
&= 1.06\\
\end{aligned}
$$
```{r}
x <- 15
μ <- 0  # mean
σ <- sqrt(200)  # standard deviation

# Calculate z-score
z <- (x - μ) / σ

# CDF CALCULATIONS
Φ <- pnorm(z)  # also, percentile of 'x'

# PDF calculation
ϕ <- dnorm(z)

cat("z-score of", x, ": ", z, "\n")
cat("pdf : (f(x) = y val of bell curve at x) = ", ϕ, "\n")
cat("cdf : Pr( X <",x, ") =", Φ, "\n")     # probability below x
cat("1 - cdf: Pr(",x, " < X ) =", (1-Φ), "\n\n")   # probability above X

# Calculate percent of data above a value 'x'
```
Probability of meeting:
```{r}
a <- -15  # lower
b <- 15  # upper

z_a <- (a - μ) / σ
z_b <- (b - μ) / σ

Φ_a <- pnorm(z_a)
Φ_b <- pnorm(z_b)

cat("Pr(",a,"< X <", b ,") :", Φ_b - Φ_a )
```
$$
\begin{aligned}
P(-1.06 < Z < 1.06) &= 0.711
\end{aligned}
$$

j) Now let m represent the number of minutes that both people agree to wait, where m can be any real number. Determine the value of m so the probability of meeting is .9.


k) Now suppose that Kareem and Kathleen can only afford to wait for 15 minutes, but they want to have at least a 90% chance of successfully meeting. Continue to assume that their arrival times follow independent normal distributions with mean 30 and the same SD as each other. Determine how small that SD needs to be in order to meet their criteria. (As always, show your work)

```{r}

```

